### 2021-02-12 Modules/Spatial_IIR-MathDerivations.tex
Sent as draft No.2

Time domain analysis of the proposed feedback based architecture, considering both propagation delay and attenuation, gives rise to
\begin{equation}
    \label{eqn:SingleSensorTemporalEquality}
    % \resizebox{.91\linewidth}{!}{
        \begin{split}
            x_{n}(t) = g\rBrace{s\rBrace{t-\tau_{pd}-\tau_{n}}
            +\sum_{m=0}^{N-1}{\alpha^{*}_{m}x_{m}\rBrace{t-\tau_{pd}-\tau_{n}}}},
        \end{split}
    % }
\end{equation}
where the first term on the right-hand-side (RHS) represents the contribution of the transmitted waveform $s(t)$ to the $n$'th array element and the second term represents the feedback contribution of the re-transmitted array signal to this same element.
Expressing the Fourier transform of \eqref{eqn:SingleSensorTemporalEquality},
\begin{equation}
    \label{eqn_singleSensorFourier}
    % \resizebox{.91\linewidth}{!}{
        \begin{split}
            X_{n} =
            g\Bigg( & \F{s}
            \exp\rBrace{-j\omega\rBrace{\tau_{pd}+\tau_{n}}}
            \\&+\sum_{m=0}^{N-1}
            {
            \alpha^{*}_{m}\omegaB\F{x}_{m}
            \exp\rBrace{-j\omega\rBrace{\tau_{pd}+\tau_{n}}}
            }\Bigg),
        \end{split}
    % }
\end{equation}
and its vector from,
$$
\F{\vx} = g\rBrace{\F{s}+\vAlphaH \F{\vx}}\vd\exp{\rBrace{-j\omega\tau\_{pd}}},
$$
we find that it can be simplified to
$$
\F{\vx} =g\rBrace{I-g\vd\vAlphaH{}e^{-j\omega\tau_{pd}}}^{-1}\vd\F{s}\exp{\rBrace{-j\omega\tau_{pd}}}.
$$
Then, denoting
\[
\phi\triangleq\omega\tau_{pd}
\]
as the round-trip signal propagation related electrical phase and using the Woodbury matrix identity \cite{woodbury1950inverting}, we find that
$$
\F{\vx}
=
\frac{    
g\vd\exp{\rBrace{-j\phi}}
}{
1 - g\aHd{}\exp{\rBrace{-j\phi}}
}\F{s}.
$$
Let $z=\vBetaH{}\vecnot{x}+\text{n}$ be the beamformer's output (see Fig.~\ref{fig:Proposed_spatialIIR_ARCH}), with Fourier transform $Z$. Considering the noiseless case $\rBrace{\text{i.e., n}=0}$, the frequency response of the FB is 
\begin{equation}
\label{eqn:GeneralFeedbackTransferFunction}
\Hba
\triangleq
\frac{\F{z}}{\F{s}} 
=
\frac{    
g\bHd{}\exp\rBrace{-j\phi}
}{
1 - g\aHd{}\exp\rBrace{-j\phi}
}.
\end{equation}
\par Note that this architecture achieves a controllable (via setting of $\vBeta$ and $\vAlpha$) and recursive (non-trivial denominator) spatial response.
As will be shown, high directivity and narrow beamwidth are obtainable by proper selection of the weights. Compared to traditional beamformers (i.e., without feedback), the performance improvement will be expressed in terms of increased aperture, narrower beamwidth and improved sidelobe attenuation.
One may observe that opposed to traditional beamformers, the array response, $\Hba,$ is not only influenced by the impinging signal DOA, for it is also range selective due to its $\phi$ dependency.
As demonstrated in Fig.~\ref{fig_rangeAzimuthSelectivity}, the combination of both angular and range selectivity enables the designer to enhance signals arriving from specific locations (grey area) rather than only specific directions.
\begin{figure}[t!]
    \begin{center}
        \begin{overpic}[width=0.65\linewidth, 
        % grid, 
        tics=10,trim=0 0 0 0]{./Media/azimuthRangSelectivity.png}
            \put (20, 23){\rotatebox{0}{\footnotesize{Angular response}}}
            \put (30.5, 47){\rotatebox{0}{\footnotesize{Enhanced radial slice}}}
        \end{overpic}
    \end{center}
    \caption{
    % A visualization of the spatial area selectivity concept.
    Combining both radial selectivity and DOA-based selectivity allows to localize the transmitter.
    }
    \label{fig_rangeAzimuthSelectivity}
\end{figure}

>>>>

Time domain analysis of the proposed feedback based architecture, considering both propagation delay and attenuation, gives rise to
\begin{equation}
    \label{eqn:SingleSensorTemporalEquality}
    % \resizebox{.91\linewidth}{!}{
        \begin{split}
            x_{n}(t) = g\rBrace{s\rBrace{t-\tau_{pd}-\tau_{n}}
            +\sum_{m=0}^{N-1}{\alpha^{*}_{m}x_{m}\rBrace{t-\tau_{pd}-\tau_{n}}}},
        \end{split}
    % }
\end{equation}
where the first term on the right-hand-side (RHS) represents the contribution of the transmitted waveform $s(t)$ to the $n$'th array element and the second term represents the feedback contribution of the re-transmitted array signal to this same element.
Expressing the Fourier transform of \eqref{eqn:SingleSensorTemporalEquality},
\begin{equation}
    \label{eqn_singleSensorFourier}
    % \resizebox{.91\linewidth}{!}{
            X_{n} =
            g\rBrace{\F{s}
            \exp\rBrace{-j\omega\rBrace{\tau_{pd}+\tau_{n}}}
            +\sum_{m=0}^{N-1}
            {
            \alpha^{*}_{m}\F{x}_{m}
            \exp\rBrace{-j\omega\rBrace{\tau_{pd}+\tau_{n}}}
            }},
    % }
\end{equation}
and its vector from,
$$
\F{\vx} = g\rBrace{\F{s}+\vAlphaH \F{\vx}}\vd\exp{\rBrace{-j\omega\tau_{pd}}},
$$
we find that it can be simplified to
$$
\F{\vx} =g\rBrace{I-g\vd\vAlphaH{}e^{-j\omega\tau_{pd}}}^{-1}\vd\F{s}\exp{\rBrace{-j\omega\tau_{pd}}}.
$$
Then, denoting
\[
\phi\triangleq\omega\tau_{pd}
\]
as the round-trip signal propagation related electrical phase.
We use the Sherman-Morrison formula \cite{sherman1950adjustment}, considered to be a result of the Woodbury matrix identity \cite{woodbury1950inverting}, stating that
\begin{equation*}
    \rBrace{I+\vecnot{u}\vecnot{v}^{T}}^{-1}=I-\frac{\vecnot{u}\vecnot{v}^{T}}{1+\vecnot{v}^{T}\vecnot{u}},
\end{equation*}
where $\vecnot{u},\vecnot{v}$ are two $N\times1$ vectors and $I$ is the identity matrix.   
Setting $\vecnot{u}=-g\vecnot{d}\exp{\rBrace{-j\omega\tau}}$ and $\vecnot{v}=\vecnot{\alpha}^{*}$ gives rise to
\begin{equation*}
    \rBrace{I+-g\vecnot{d}\vecnot{\alpha}^{H}\exp{\rBrace{-j\omega\tau}}}^{-1} 
    = I+\frac{g\vecnot{d}\vecnot{\alpha}^{H}\exp{\rBrace{-j\omega\tau}}}{1-g\vecnot{\alpha}^{H}\vecnot{d}\exp{\rBrace{-j\omega\tau}}}
\end{equation*}
which results in
\begin{equation*}
    \frac{\rBrace{I-g\vecnot{\alpha}^{H}\vecnot{d}\exp{\rBrace{-j\omega\tau}}+g\vecnot{d}\vecnot{\alpha}^{H}\exp{\rBrace{-j\omega\tau}}}\vecnot{d}}{1-g\vecnot{\alpha}^{H}\vecnot{d}\exp{\rBrace{-j\omega\tau}}}.
\end{equation*}
Obviously the numerator's 2nd and 3rd terms vanish using the commutativity concept stating that
\begin{equation*}
    \vecnot{\alpha}^{H}\vecnot{d}\vecnot{d} = \vecnot{d}\vecnot{\alpha}^{H}\vecnot{d}
\end{equation*}
which finally results in
$$
\F{\vx}
=
\frac{    
g\vd\exp{\rBrace{-j\phi}}
}{
1 - g\aHd{}\exp{\rBrace{-j\phi}}
}\F{s}.
$$
Let $z=\vBetaH{}\vecnot{x}+\text{n}$ be the beamformer's output (see Fig.~\ref{fig:Proposed_spatialIIR_ARCH}), with Fourier transform $Z$. Considering the noiseless case $\rBrace{\text{i.e., n}=0}$, the frequency response of the FB is 
\begin{equation}
\label{eqn:GeneralFeedbackTransferFunction}
\Hba
\triangleq
\frac{\F{z}}{\F{s}} 
=
\frac{    
g\bHd{}\exp\rBrace{-j\phi}
}{
1 - g\aHd{}\exp\rBrace{-j\phi}
}.
\end{equation}
\par Note that this architecture achieves a controllable (via setting of $\vBeta$ and $\vAlpha$) and recursive (non-trivial denominator) spatial response.
As will be shown, high directivity and narrow beamwidth are obtainable by a proper selection of the weights. Compared to traditional beamformers (i.e., without feedback), the performance improvement will be expressed in terms of increased aperture, narrower beamwidth and improved sidelobe attenuation.
One may observe that opposed to traditional beamformers, the array response, $\Hba,$ is not only influenced by the impinging signal DOA, since it is also range selective due to its $\phi$ dependency.
As demonstrated in Fig.~\ref{fig_rangeAzimuthSelectivity}, the combination of both angular and range selectivity enables the designer to enhance signals arriving from specific locations (grey area) rather than only specific directions.
\begin{figure}[t!]
    \begin{center}
        \begin{overpic}[width=0.55\linewidth, 
        % grid, 
        tics=10,trim=0 0 0 0]{./Media/azimuthRangSelectivity.png}
            \put (20, 23){\rotatebox{0}{\footnotesize{Angular response}}}
            \put (32, 47){\rotatebox{0}{\footnotesize{Enhanced radial slice}}}
        \end{overpic}
    \end{center}
    \caption{
    % A visualization of the spatial area selectivity concept.
    Combining both radial selectivity and DOA-based selectivity allows to localize the transmitter.
    }
    \label{fig_rangeAzimuthSelectivity}
\end{figure}

### 2021-02-12 Modules/abstractHebrew.tex
Sent as draft No.2

עם התקדמות המדע, רמת הדיוק הנדרשת בשיערוך עלתה ובכדי לעקוף את הגבולות התיאורטיים (פורייה למשל), עלה הרעיון להלחמת מידע ממספר מדידות בו זמנית.

@@

עדויות לכך ניתן למצוא עוד בתקופת מלחמת העולם השניה.

@@

במערכות תקשורת, הצורך בסינון מרחבי הוא מן המעלה הראשונה, שכן במציאות מרובת המשדרים והמקלטים, במידה ולא היה שימוש במערכים, הפרעות הדדיות היו מונעות את פענוח האות הנקלט. במקביל, לא רק סינון אלא גם הכוונת אלומת השידור הפכה לאפשרית בזכות השימוש בו-זמנית במספר רב של משדרים

@@

לשם המחשת הפתרון שמוצע, עבודה זו מתמקדת בתחום האיכון - בסביבה אקוסטית או אלקטרו מגנית.

@@

על פניו, מערכת דומה למכ"ם, השולח אות ומשערך את מיקומי מטרות על פי האות החוזר.
החידוש בהצעה נעוץ בכך שהאות המשודר אינו רק אות מלאכותי הנוצר ע"י המערכת אלא משולב בו האות המסונן הנקלט ע"י המערך.

>>>>

עם התקדמות המדע, רמת הדיוק הנדרשת בשיערוך עלתה ובכדי לעקוף את הגבולות התיאורטיים (פורייה למשל), עלה הרעיון לעיבוד מקבילי של מידע ממספר רגשים.

@@

עדויות לכך ניתן למצוא עוד בתקופת מלחמת העולם השניה בדמות רדאר ה
\textenglish{Mammut}
שנבנה על ידי הגרמנים לזיהוי מוקדם של מטרות טסות בגבהים עד 8 ק"מ ומרחקים של עד 300 ק"מ.

@@

במערכות תקשורת, הצורך בסינון מרחבי הוא מן המעלה הראשונה, שכן במציאות מרובת המשדרים והמקלטים, במידה ולא היה שימוש במערכים, הפרעות הדדיות היו מונעות את פענוח האות הנקלט. במקביל, לא רק סינון אלא גם הכוונת אלומת השידור הפכה לאפשרית בזכות השימוש במערכי מופע

@@

למשל, עלתה הצעה לשערך את השהיית האות בין רגשים שכנים, וע"י כך לייצר היזון חוזר בצורה מלאכותית.
גישה זו חשופה לשגיאות מדידה וחוטאת למטרת העיבוד המרחבי הטהור - שכן היא מערבת עיבוד זמני.
הצעה נוספת היתה להתייחס לתת מערכים (מתוך מערך אחד) חופפים כאל הסחות זמניות של מערך אחד קטן יותר ובכך לייצר קירוב סופי לתגובת ההלם האינסופית.
הצעה זו אינה אלא פרשנות שונה לעיבוד מבוסס 
שכן בסופו של תהליך העיבוד הינו לינארי ומכיל מספר סופי של דגמים במערכת בכל זמן נתון.
שיטה מעניינת, המהווה בסיס למחקר ענף, מטילה את הדגימה המרחבית והדגימה הזמנית אל מישור התדר, בו גל מישורי מיוצג על ידי קו ישר המוטה לפי זוית ההגעה של הגל הפוגע.
בהמשך לכך, מסננים במרחב התדר הדו מימדי מתוכננים על מנת להגביר אותות שמגיעים מכיוונים רצויים.
גישה זו בשימוש רחב בתחום החקר הסיסמי וסובלת בעיקר מחוסר היכולת לתכנן מסננים מדויקים היות והמסננים מחזוריים בתדר, עובדה הגורמת לעיוותים בקצוות מרחב התדר.
עוד על כן, גם שיטה זו אינה סינון מרחבי טהור ולכן אינה מהווה פתרון לשאלת המחקר אותה ניסינו לפתור.

@@

על פניו, מערכת דומה למכ"ם, הפולט אות ומשערך את מיקומי מטרות על פי האות החוזר.
החידוש בהצעה נעוץ בכך שהאות המשודר אינו רק אות מלאכותי הנוצר ע"י המערכת אלא משולב בו האות המסונן הנקלט ע"י המערך, בכך נוצר היזון חוזר בין מהמערך והמטרה.

### 2021-02-12 Modules/arrayPerformance_beamwidth_LOCAL.tex
Sent as draft No.2

The half power beamwidth (HPBW), denoted as $\dThetaHPBW$ quantifies the array's main lobe narrowness.
It represents the DOA where the beampattern's energy reduces to half of its maximal value.
For standard ULA, assuming large $N$ values, it is known \cite{van2004optimum} that
$$\dThetaHPBW/2= 1.4/N.$$
\par In App.~\ref{apdx_HPBW} we extend this known result for any $r\geq 0$. It turns out that for large $N$, the HPBW is obtained by solving for $x$ the equality
\begin{equation}\label{eq_HPBW}
    % \resizebox{1\linewidth}{!}{
        \begin{split}
            \rBrace{r^{2}-4r+2}\frac{\sin{\rBrace{x}}^{2}}{x^{2}}+r\frac{\sin{\rBrace{2x}}}{x}-1=0
        \end{split}
    % }
\end{equation}
where we define $x\triangleq{} N\dTheta_{HPBW}/2$. In Fig.~\ref{fig_feedbackULA_HPBW_Nx_vs_N_variousR} we plot the numerical solution of \eqref{eq_HPBW} for various values of $r$ and $N$, showing that $x$ reaches its limit around $N=20$. Also note that for $r=0$ we obtain the known result of standard ULA with the limiting factor of $1.4$.
Having the limiting factors for various values of the gain mismatch $r$, we investigate the feedback related improvement and express the HPBW by
\[
\dThetaHPBW/2\approx \frac{1.4}{f(r)N}.
\]
Note that $f(r)$ represents the array aperture improvement factor, compared to the standard ULA. 
Numerical evaluation (see Fig.~\ref{fig_feedbackULA_beamwidth_limit_r_dependent}) shows that $f(r)$ may be approximated with a second order polynomial
\begin{equation}
    \label{eq_Bapprox}
    f\rBrace{r}\approx\frac{1.4}{\rBrace{1-r}\rBrace{-0.4r+1.4}}.
\end{equation}
Note that for an accurate gain match (i.e., $r\to1$), the RHS of \eqref{eq_Bapprox} tends towards infinity, implying that the equivalent array has an infinite number  of elements ($f\rBrace{r}N$), hence obtaining perfect spatial selectivity.
\begin{figure}[t]
    \begin{center}
        \begin{overpic}[width=0.65\linewidth, 
        %grid, 
        tics=10,trim=0 0 0 0]{./Media/spatial_IIR_MATLAB/arrayParameters/HPBW_vs_N_various_r.eps}
            \put (4, 75){\footnotesize{$N\dThetaHPBW/2$}}
            \put (50, 62.5) {\footnotesize{$r=0$}}
            \put (50, 54) {\footnotesize{$r=0.1$}}
            \put (50, 39.5) {\footnotesize{$r=0.3$}}
            \put (50, 28.5) {\footnotesize{$r=0.5$}}
            \put (50, 19.75) {\footnotesize{$r=0.7$}}
            \put (50, 12.5) {\footnotesize{$r=0.9$}}
            \put (50, 2) {\footnotesize{$N$}}
        \end{overpic}
    \end{center}
     \caption{Plot of $x=N\dThetaHPBW/2$ vs. $N$, for various $r$ values, obtained by numerically solving \eqref{eq_HPBW}.}
    \label{fig_feedbackULA_HPBW_Nx_vs_N_variousR}
\end{figure}
\begin{figure}[t]
    \begin{center}
        \begin{overpic}[width=0.65\linewidth, 
        % grid, 
        tics=10,trim=0 0 0 0]{./Media/HPBW_limit_vs_r.eps}
            \put (39.5, 63.5) {\scriptsize{Numerical solution of \eqref{eq_HPBW}}}
            \put (39.5, 58.25) {\scriptsize{Polynomial fitting \eqref{eq_Bapprox}}}
            \put (39.5, 52.5) {\footnotesize{$\log_{10}f\rBrace{r}$}}
            \put (85, 75) {\footnotesize{$\log_{10}f\rBrace{r}$}}
            \put (4, 75){\footnotesize{$N\dThetaHPBW/2$}}
            \put (50, 2) {\footnotesize{$r$}}
        \end{overpic}
    \end{center}
    \caption{Evaluation of $N\dThetaHPBW/2$ for $N=100$ and its approximation $1.4/f\rBrace{r}$  (marked by red diamonds). $f\rBrace{r}$ is also presented, in logarithmic scale (dotted curve).} 
    \label{fig_feedbackULA_beamwidth_limit_r_dependent}
\end{figure}

>>>>

% The half power beamwidth (HPBW), denoted as $\dThetaHPBW$ quantifies the array's main lobe narrowness.
% It represents the DOA where the beampattern's energy reduces to half of its maximal value.
\par In App.~\ref{apdx_HPBW} we extend the known result of \eqref{eq_known_HBPW} for any $r\geq 0$. It turns out that for large $N$, the HPBW is obtained by solving for $x$ the equality
\begin{equation}\label{eq_HPBW}
    % \resizebox{1\linewidth}{!}{
        \begin{split}
            \rBrace{r^{2}-4r+2}\frac{\sin{\rBrace{x}}^{2}}{x^{2}}+r\frac{\sin{\rBrace{2x}}}{x}-1=0
        \end{split}
    % }
\end{equation}
where we define $x\triangleq{} N\dTheta_{HPBW}/2$. In Fig.~\ref{fig_feedbackULA_HPBW_Nx_vs_N_variousR} we plot the numerical solution of \eqref{eq_HPBW} for various values of $r$ and $N$, showing that $x$ reaches its limit around $N=20$. Also note that for $r=0$ we obtain the known result of standard ULA with the limiting factor of $1.4$.
Having the limiting factors for various values of the gain mismatch $r$, we investigate the feedback related improvement and express the HPBW by
\[
\dThetaHPBW/2\approx \frac{1.4}{f(r)N}.
\]
Note that $f(r)$ represents the array aperture improvement factor, compared to the standard ULA.
To find a suitable expression for $f\rBrace{r}$, we computed $\Hr_{\Delta\theta,\Delta\phi=0,r}$ using $N=100$ on a fine grid of $r$ values (actually slicing a vertical column of Fig.~\ref{fig_feedbackULA_HPBW_Nx_vs_N_variousR} at $N=100$) and looked for proper polynomial approximation using MATLAB\textsuperscript{\textregistered}.
It turns out that the 2-nd order polynomial best suits the computation and that the best fitted expression is 
\begin{equation}
    \label{eq_Bapprox}
    f\rBrace{r}\approx\frac{1.4}{\rBrace{1-r}\rBrace{-0.4r+1.4}}.
\end{equation}
Note that for an accurate gain match (i.e., $r\to1$), the RHS of \eqref{eq_Bapprox} tends towards infinity, implying that the equivalent array has an infinite number  of elements ($f\rBrace{r}N$), hence obtaining perfect spatial selectivity.
\begin{figure}[t]
    \begin{center}
        \begin{overpic}[width=0.65\linewidth, 
        %grid, 
        tics=10,trim=0 0 0 0]{./Media/spatial_IIR_MATLAB/arrayParameters/HPBW_vs_N_various_r.eps}
            \put (4, 75){\footnotesize{$N\dThetaHPBW/2$}}
            \put (50, 62.5) {\footnotesize{$r=0$}}
            \put (50, 54) {\footnotesize{$r=0.1$}}
            \put (50, 39.5) {\footnotesize{$r=0.3$}}
            \put (50, 28.5) {\footnotesize{$r=0.5$}}
            \put (50, 19.75) {\footnotesize{$r=0.7$}}
            \put (50, 12.5) {\footnotesize{$r=0.9$}}
            \put (50, 2) {\footnotesize{$N$}}
        \end{overpic}
    \end{center}
     \caption{Plot of $x=N\dThetaHPBW/2$ vs. $N$, for various $r$ values, obtained by numerically solving \eqref{eq_HPBW}.}
    \label{fig_feedbackULA_HPBW_Nx_vs_N_variousR}
\end{figure}
\begin{figure}[t]
    \begin{center}
        \begin{overpic}[width=0.65\linewidth, 
        % grid, 
        tics=10,trim=0 0 0 0]{./Media/HPBW_limit_vs_r.eps}
            \put (39.5, 63.5) {\scriptsize{Numerical solution of \eqref{eq_HPBW}}}
            \put (39.5, 58.25) {\scriptsize{Polynomial fitting \eqref{eq_Bapprox}}}
            \put (39.5, 52.5) {\footnotesize{$\log_{10}f\rBrace{r}$}}
            \put (85, 75) {\footnotesize{$\log_{10}f\rBrace{r}$}}
            \put (4, 75){\footnotesize{$N\dThetaHPBW/2$}}
            \put (50, 2) {\footnotesize{$r$}}
        \end{overpic}
    \end{center}
    \caption{Evaluation of $N\dThetaHPBW/2$ for $N=100$ and its approximation $1.4/f\rBrace{r}$  (marked by red diamonds). $f\rBrace{r}$ is also presented, in logarithmic scale (dotted curve).} 
    \label{fig_feedbackULA_beamwidth_limit_r_dependent}
\end{figure}
